### 2017-08-28 sections/conclusion.tex
Apply most suggestions of Reviewer #3 except ones dealing with diagrams in Section Results
Fix abbreviations (FEM and SVD)

The goal of this paper is to present the use of SVD decomposition in compression of results from the Finite element method. Implementation details of the compression algorithm have been provided, and the results of its application on real data have been presented. The implemented algorithm is able to compress arbitrary data using low-rank approximation matrices. When the maximum allowed error was set to $10^{-5}$, the compression ratio was at most 10\% for all tested results. In many cases compression ratio can be even better -- bellow 1\% of the original size. The important property of the compression algorithm is the fact that the approximation error can be set in advance and there is a guarantee that it will not be exceeded.

The main disadvantage is the computational complexity of the compression algorithm. SVD decomposition is a very time-consuming operation. However, this operation is performed only once after FEM analysis is finished and before the post-processing is started. Also, the randomized version of the SVD decomposition algorithm is much faster and can be used if a slight increase of approximation error is tolerated.

Another complication is the necessity to use a special format to store compressed data in the form of SVD decomposition, and the post-processor must be updated to use this new format. Post-processor must perform matrix multiplication to get the original results. However, in a usual case the data for one analysis step are needed at a time, i.e. multiplication of a single row is enough. Detailed description of the post-processor implementation is beyond the scope of this paper.

>>>>

The goal of this paper is to present the use of the SVD in compression of results from the Finite Element Method. Implementation details of the compression algorithm have been provided, and the results of its application on real data have been presented. The implemented algorithm is able to compress arbitrary data using low-rank approximation matrices. When the maximum allowed error was set to $10^{-5}$, the compression ratio was at most 10\% for all tested results. In many cases compression ratio can be even better -- bellow 1\% of the original size. The important property of the compression algorithm is the fact that the approximation error can be set in advance and there is a guarantee that it will not be exceeded.

The main disadvantage is the computational complexity of the compression algorithm. SVD is a very time-consuming operation. However, this operation is performed only once after FEM analysis is finished and before the post-processing is started. Also, the randomized version of the decomposition algorithm is much faster and can be used if a slight increase of approximation error is tolerated.

Another complication is the necessity to use a special format to store compressed data in the form of matrix decompositions, and the post-processor must be updated to use this new format. Post-processor must perform matrix multiplication to get the original results. However, in a usual case the data for one analysis step are needed at a time, i.e. multiplication of a single row is enough. Detailed description of the post-processor implementation is beyond the scope of this paper.

### 2017-08-28 sections/implementation.tex
Fix abbreviations (FEM and SVD)

SVD decomposition is applied on matrices. The first thing to do is therefore to assemble an input matrix. Results from the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements. As an example, temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc. can serve. There are two similar sets of results. One is generated by a non-linear algorithms, where several incremental steps are stored and the other is generated by time integration, where results in particular time steps are stored. In order to compress data, an auxiliary matrix $\mtrx{A}$ has to be assembled. The number of rows of the matrix $\mtrx{A}$ is equal to the number of incremental or time steps while the number of columns is equal to the number of points in which the results are stored. Such auxiliary matrix is assembled for each scalar field and for each component of the vector and tensor fields. It means, three matrices corresponding to the displacement in the $x$, $y$ and $z$ directions are assembled for the vector of displacements in three-dimensional problems.

@@

\subsection{Algorithm description}
Once the SVD decomposition is calculated, the compression algorithm removes a certain number of singular values and corresponding singular vectors. The remaining singular values and vectors represent the compressed data. There are two strategies that influence the way how to preserve the number of singular values -- resulting size and quality. Each strategy is assigned a control parameter that determines compression ratio or approximation error.

@@

Computational complexity of the exact SVD decomposition algorithm is $\mathrm{O}(m^2n)$, where $m<n$. This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of the execution time on the varying matrix dimension is shown. The results of the benchmarks are depicted in Figure \ref{fig:ExeTime_rows} and Figure \ref{fig:ExeTime_columns}. Several observations were made from the results:

@@

Storage size of the SVD decomposition itself can also be optimized. $\mtrx{S}$, being a diagonal matrix, can be stored as single list of singular values $s_{i}$, or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.

>>>>

SVD is applied on matrices. The first thing to do is therefore to assemble an input matrix. Results from the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements. As an example, temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc. can serve. There are two similar sets of results. One is generated by a non-linear algorithms, where several incremental steps are stored and the other is generated by time integration, where results in particular time steps are stored. In order to compress data, an auxiliary matrix $\mtrx{A}$ has to be assembled. The number of rows of the matrix $\mtrx{A}$ is equal to the number of incremental or time steps while the number of columns is equal to the number of points in which the results are stored. Such auxiliary matrix is assembled for each scalar field and for each component of the vector and tensor fields. It means, three matrices corresponding to the displacement in the $x$, $y$ and $z$ directions are assembled for the vector of displacements in three-dimensional problems.

@@

\subsection{Algorithm description}
Once the SVD is calculated, the compression algorithm removes a certain number of singular values and corresponding singular vectors. The remaining singular values and vectors represent the compressed data. There are two strategies that influence the way how to preserve the number of singular values -- resulting size and quality. Each strategy is assigned a control parameter that determines compression ratio or approximation error.

@@

Computational complexity of the exact SVD algorithm is $\mathrm{O}(m^2n)$, where $m<n$. This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of the execution time on the varying matrix dimension is shown. The results of the benchmarks are depicted in Figure \ref{fig:ExeTime_rows} and Figure \ref{fig:ExeTime_columns}. Several observations were made from the results:

@@

Storage size of the SVD itself can also be optimized. $\mtrx{S}$, being a diagonal matrix, can be stored as single list of singular values $s_{i}$, or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.

### 2017-08-28 sections/introduction.tex
Fix abbreviations (FEM and SVD)

Amount of data produced by a complex finite element analysis can be enormous and typical personal computer is not capable to store, process and visualize the results in reasonable time. Singular Value Decomposition (SVD) is a well known factorization method that provides rich information about matrix systems. One of its many applications is the image compression where it can significantly reduce the size of the data representing an image while preserving the quality of image appearance.

@@

Compression methods usually yield approximated data. In the following text, the term \textit{approximation error} denotes an error resulted from compression, i.e. difference between original results of FEM analysis and their compressed form. It should not be confused with the error of the Finite element method itself that yields approximate solution to mathematical problems used to model physical reality.

>>>>

Amount of data produced by a complex finite element analysis can be enormous and typical personal computer is not capable to store, process and visualize the results in reasonable time. Singular Value Decomposition is a well known factorization method that provides rich information about matrix systems. One of its many applications is the image compression where it can significantly reduce the size of the data representing an image while preserving the quality of image appearance.

@@

Compression methods usually yield approximated data. In the following text, the term \textit{approximation error} denotes an error resulted from compression, i.e. difference between original results of FEM analysis and their compressed form. It should not be confused with the error of the Finite Element Method itself that yields approximate solution to mathematical problems used to model physical reality.

### 2017-08-28 sections/math.tex
Fix abbreviations (FEM and SVD)

Singular value decomposition (SVD) \cite{Baker2005, Kalman1996, Golub1996} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
matrix $\mtrx{S} \in \mathbb{R}^{m \times n}$, and the transpose of an orthogonal matrix $\mtrx{V} \in \mathbb{R}^{n \times n}$:

@@

\noindent
The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of the left and the right singular vectors of matrix $\mtrx{A}$, and a nonnegative diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$. For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{r})$ floating-point operations, substantially less than classical algorithms.

>>>>

Singular value decomposition \cite{Baker2005, Kalman1996, Golub1996} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
matrix $\mtrx{S} \in \mathbb{R}^{m \times n}$, and the transpose of an orthogonal matrix $\mtrx{V} \in \mathbb{R}^{n \times n}$:

@@

\noindent
The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of the left and the right singular vectors of matrix $\mtrx{A}$, and a nonnegative diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$. For a dense input matrix, randomized SVD algorithm requires $\mathrm{O}(mn \log{r})$ floating-point operations, substantially less than classical algorithms.

### 2017-08-28 sections/results.tex
Minor fix (add comma)
Fix abbreviations (FEM and SVD)

The first benchmark was designed to measure computational complexity of the SVD decomposition algorithm. Series of 100 random matrices with standard distribution were generated and execution times were recorded and averaged. The execution time with respect to the number of the stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the execution time with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}. Especially in Figure \ref{fig:ExeTime_columns}, it is clearly visible that the randomized SVD algorithm is much faster than the classical one.

@@

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
\caption{Variation of execution time of standard and randomized SVD decompositions calculated for reactor containment analysis results.}
\label{fig:temelin:ExeTime}
\end{figure}

The memory consumption of compressed results for reactor containment analysis is summarized in Table \ref{tab:mem-consum}. For different values of compression ratio it shows memory size in megabytes. In this benchmark the compression ratio $c$ is an input parameter to the compression algorithm. As follows from the Equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD decomposition. Size factor describes the final outcome of compression when compared to the original size.

>>>>

The first benchmark was designed to measure computational complexity of the SVD algorithm. Series of 100 random matrices with standard distribution were generated and execution times were recorded and averaged. The execution time with respect to the number of the stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the execution time with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}. Especially in Figure \ref{fig:ExeTime_columns}, it is clearly visible that the randomized SVD algorithm is much faster than the classical one.

@@

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
\caption{Variation of execution time of standard and randomized SVD calculated for reactor containment analysis results.}
\label{fig:temelin:ExeTime}
\end{figure}

The memory consumption of compressed results for reactor containment analysis is summarized in Table \ref{tab:mem-consum}. For different values of compression ratio it shows memory size in megabytes. In this benchmark, the compression ratio $c$ is an input parameter to the compression algorithm. As follows from the Equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD. Size factor describes the final outcome of compression when compared to the original size.
